---
title: 'Story04:How much do we get paid?'
author: "Mahmud Hasan Al Raji"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
We have used the term "Data Practitioner" as a generic job descriptor because we have so many different job role titles for individuals whose work activities overlap including Data Scientist, Data Engineer, Data Analyst, Business Analyst, Data Architect, etc.In this assignment we will answer the question, "How much do we get paid?" The analysis and data visualizations will address the variation in average salary based on role descriptor and state.

It is noted that the annual salary data (up to October 2023) for Data Scientists, Data Engineers, Data Analysts, and Business Analysts role by US states are collected from ZipRecruiter, which is an online employment marketplace and job search engine that connects employers and job seekers. The data, originally presented in a tabular format on ZipRecruiter, were converted into CSV file format and subsequently uploaded to my GitHub repository for further analysis. Data links are given in Data Reference section.  


## Load library
```{r error=FALSE}
#library(readxl)
library(tidyverse)
#library(rvest)
#library(xml2)
```


## Get Data
```{r}
data_scientist<-read.csv("https://raw.githubusercontent.com/Raji030/story4_data/main/data_scientist.csv")
data_engineer<-read.csv("https://raw.githubusercontent.com/Raji030/story4_data/main/data_engineer.csv")
data_analyst<-read.csv("https://raw.githubusercontent.com/Raji030/story4_data/main/data_analyst.csv")
business_analyst<-read.csv("https://raw.githubusercontent.com/Raji030/story4_data/main/data_business_analyst.csv")
```


## Data Preparation
```{r}

## For Data Scientist: 

# Remove "$" and ","
data_scientist$Annual.Salary <- as.numeric(gsub("[$,]", "", data_scientist$Annual.Salary))
data_scientist$Hourly.Wage <- as.numeric(gsub("[$]", "", data_scientist$Hourly.Wage))

# Rename Annual.Salary column to Annual_Salary, and Hourly.Wage to Hourly_Wage
colnames(data_scientist)[colnames(data_scientist) == "Annual.Salary"] <- "Annual_Salary"
colnames(data_scientist)[colnames(data_scientist) == "Hourly.Wage"] <- "Hourly_Wage"

# Remove Monthly.Pay and Weekly.Pay columns
data_scientist <- data_scientist[, !(colnames(data_scientist) %in% c("Monthly.Pay", "Weekly.Pay"))]

# Arrange "State" column in ascending order
data_scientist <- data_scientist[order(data_scientist$State), ]

# Include Role column
data_scientist<-data_scientist%>%mutate(Role="Data Scientist")

head(data_scientist)

## For Data Engineer:

# Remove "$" and ","
data_engineer$Annual.Salary <- as.numeric(gsub("[$,]", "", data_engineer$Annual.Salary))
data_engineer$Hourly.Wage <- as.numeric(gsub("[$]", "", data_engineer$Hourly.Wage))

# Rename Annual.Salary column to Annual_Salary, and Hourly.Wage to Hourly_Wage
colnames(data_engineer)[colnames(data_engineer) == "Annual.Salary"] <- "Annual_Salary"
colnames(data_engineer)[colnames(data_engineer) == "Hourly.Wage"] <- "Hourly_Wage"

# Remove Monthly.Pay and Weekly.Pay columns
data_engineer<- data_engineer[, !(colnames(data_engineer) %in% c("Monthly.Pay", "Weekly.Pay"))]

# Arrange "State" column in ascending order
data_engineer <- data_engineer[order(data_engineer$State), ]

# Include Role column
data_engineer<-data_engineer%>%mutate(Role="Data Engineer")

head(data_engineer)

## For Data Analyst:

# Remove "$" and ","
data_analyst$Annual.Salary <- as.numeric(gsub("[$,]", "", data_analyst$Annual.Salary))
data_analyst$Hourly.Wage <- as.numeric(gsub("[$]", "", data_analyst$Hourly.Wage))

# Rename Annual.Salary column to Annual_Salary, and Hourly.Wage to Hourly_Wage
colnames(data_analyst)[colnames(data_analyst) == "Annual.Salary"] <- "Annual_Salary"
colnames(data_analyst)[colnames(data_analyst) == "Hourly.Wage"] <- "Hourly_Wage"

# Remove Monthly.Pay and Weekly.Pay columns
data_analyst<- data_analyst[, !(colnames(data_analyst) %in% c("Monthly.Pay", "Weekly.Pay"))]

# Arrange "State" column in ascending order
data_analyst <- data_analyst[order(data_analyst$State), ]

# Include Role column
data_analyst<-data_analyst%>%mutate(Role="Data Analyst")
head(data_analyst)

## For Business Analyst:

# Remove "$" and ","
business_analyst$Annual.Salary <- as.numeric(gsub("[$,]", "", business_analyst$Annual.Salary))
business_analyst$Hourly.Wage <- as.numeric(gsub("[$]", "", business_analyst$Hourly.Wage))

# Rename Annual.Salary column to Annual_Salary, and Hourly.Wage to Hourly_Wage
colnames(business_analyst)[colnames(business_analyst) == "Annual.Salary"] <- "Annual_Salary"
colnames(business_analyst)[colnames(business_analyst) == "Hourly.Wage"] <- "Hourly_Wage"

# Remove Monthly.Pay and Weekly.Pay columns
business_analyst<- business_analyst[, !(colnames(business_analyst) %in% c("Monthly.Pay", "Weekly.Pay"))]

# Arrange "State" column in ascending order
business_analyst <- business_analyst[order(business_analyst$State), ]

# Include Role column
business_analyst<-business_analyst%>%mutate(Role="Business Analyst")
head(business_analyst)

```


```{r}
## For Data Scientist:
# Find maximum, minimum, and average annual salary
max_salary <- max(data_scientist$Annual_Salary)
min_salary <- min(data_scientist$Annual_Salary)
avg_salary1 <- mean(data_scientist$Annual_Salary)

# Find state with maximum salary
states_max_salary <- data_scientist$State[data_scientist$Annual_Salary == max_salary]

# Find state with minimum salary
states_min_salary <- data_scientist$State[data_scientist$Annual_Salary == min_salary]

# Find number of states with a salary greater than annual average
states_above_avg <- sum(data_scientist$Annual_Salary > avg_salary1)

# Show results
cat("Maximum Annual Salary:", max_salary, "in", states_max_salary, "\n")
cat("Minimum Annual Salary:", min_salary, "in", states_min_salary, "\n")
cat("Average Annual Salary:", avg_salary1, "\n")
cat("Number of states that get annual salary greater than average for Data Scientists:", states_above_avg, "\n")

## For Data Engineer:
# Find maximum, minimum, and average annual salary
max_salary <- max(data_engineer$Annual_Salary)
min_salary <- min(data_engineer$Annual_Salary)
avg_salary2 <- mean(data_engineer$Annual_Salary)

# Find state with maximum salary
states_max_salary <- data_engineer$State[data_engineer$Annual_Salary == max_salary]

# Find state with minimum salary
states_min_salary <- data_engineer$State[data_engineer$Annual_Salary == min_salary]

# Find number of states with a salary greater than annual average
states_above_avg <- sum(data_engineer$Annual_Salary > avg_salary2)

# Show results
cat("Maximum Annual Salary:", max_salary, "in", states_max_salary, "\n")
cat("Minimum Annual Salary:", min_salary, "in", states_min_salary, "\n")
cat("Average Annual Salary:", avg_salary2, "\n")
cat("Number of states that get annual salary greater than average for Data Engineers:", states_above_avg, "\n")

# For Data Analyst:
# Find maximum, minimum, and average annual salary
max_salary <- max(data_analyst$Annual_Salary)
min_salary <- min(data_analyst$Annual_Salary)
avg_salary3 <- mean(data_analyst$Annual_Salary)

# Find state with maximum salary
states_max_salary <- data_analyst$State[data_analyst$Annual_Salary == max_salary]

# Find state with minimum salary
states_min_salary <- data_analyst$State[data_analyst$Annual_Salary == min_salary]

# Find number of states with a salary greater than annual average
states_above_avg <- sum(data_analyst$Annual_Salary > avg_salary3)

# Show results
cat("Maximum Annual Salary:", max_salary, "in", states_max_salary, "\n")
cat("Minimum Annual Salary:", min_salary, "in", states_min_salary, "\n")
cat("Average Annual Salary:", avg_salary3, "\n")
cat("Number of states that get annual salary greater than average for Data Analysts:", states_above_avg, "\n")

# For Business Analyst:

# Find maximum, minimum, and average annual salary
max_salary <- max(business_analyst$Annual_Salary)
min_salary <- min(business_analyst$Annual_Salary)
avg_salary4 <- mean(business_analyst$Annual_Salary)

# Find state with maximum salary
states_max_salary <- business_analyst$State[business_analyst$Annual_Salary == max_salary]

# Find state with minimum salary
states_min_salary <- business_analyst$State[business_analyst$Annual_Salary == min_salary]

# Find number of states with a salary greater than annual average
states_above_avg <- sum(business_analyst$Annual_Salary > avg_salary4)

# Show results
cat("Maximum Annual Salary:", max_salary, "in", states_max_salary, "\n")
cat("Minimum Annual Salary:", min_salary, "in", states_min_salary, "\n")
cat("Average Annual Salary:", avg_salary4, "\n")
cat("Number of states that get annual salary greater than average for Business Analysts:", states_above_avg, "\n")

# Create a dataframe for Data Scientist, Data Engineer, Data Analyst, and Business Analyst
role <- c("Data Scientist", "Data Engineer", "Data Analyst", "Business Analyst")
annual_avg_salary <- c(avg_salary1, avg_salary2, avg_salary3,avg_salary4)

df_new <- data.frame(Role = role, Annual_Avg_Salary = annual_avg_salary)
print(df_new)

```


## Data Visualizations:
```{r}
# Increase the height of the plot
#height <- 12  # Specify the height of the plot
#width <- 20  # Specify the width of the plot
#options(repr.plot.height=height, repr.plot.width=width)

# Create bar plot
p<-ggplot(data_scientist, aes(x = State, y = Annual_Salary, fill ="Data Scientist", label = State)) +
  geom_bar(stat = "identity") +
  #geom_text(position = position_stack(vjust =0.5),angle = 90, hjust =0.5) +
  labs(title = "Annual Salary for Data Scientist role in US States",
       x = "State", y = "Annual Salary(in dollars)", fill = "Role") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5, margin = margin(b =10)))  # Adjust the margin)
  
# Set breaks for y-axis
p<-p + scale_y_continuous(breaks = seq(0, max(data_scientist$Annual_Salary), by=10000))

# Add a line for the average annual salary
p <- p + geom_segment(aes(x = 0, xend =50.5, y = avg_salary1, yend = avg_salary1), color = "black", linewidth = 0.001)

#Label average annual salary line
p <- p + annotate("text", x = 35, y = avg_salary1+3500, label = paste("Avg Salary Line"), 
                  vjust = -1, hjust = -0.1, color = "black", size = 3)
# Add annotations
p + annotate("text", x=35, y = 200000, label = "(Maximum Annual Salary: $145027 in New York,", size=3.3)+ 
  annotate("text", x =35, y = 190000, label = "Minimum Annual Salary: $96232 in Arkansas,",size=3.3)+
  annotate("text",x=35,y=180000, label="Average Annual Salary: $116193.76,",size=3.3 )+
  annotate("text",x=35,y=170000, label="26 states get more than annual average salary)",size=3.3 )
```

```{r}
# Create bar plot
p <- ggplot(data_engineer, aes(x = State, y = Annual_Salary, fill = "Data Engineer", label = State)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = "orange", name = "Role", labels = "Data Engineer") +
  labs(title = "Annual Salary for Data Engineer role in US States",
       x = "State", y = "Annual Salary (in dollars)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5, margin = margin(b = 10)))  # Adjust margin
# Set breaks for y-axis
p<-p + scale_y_continuous(breaks = seq(0, max(data_engineer$Annual_Salary), by=10000))

# Add a line for the average annual salary
p <- p + geom_segment(aes(x = 0, xend =50.5, y = avg_salary2, yend = avg_salary2), color = "black", linewidth = 0.001)

#Label average annual salary line
p <- p + annotate("text", x =13, y = avg_salary2+2500, label = paste("Avg Salary Line"), 
                  vjust = -1, hjust = -0.1, color = "black", size = 3)
# Add annotations
p + annotate("text", x=35, y = 200000, label = "(Maximum Annual Salary: $149976 in Nevada ,", size=3.3)+ 
  annotate("text", x =35, y = 190000, label = "Minimum Annual Salary: $96743 in Florida ,",size=3.3)+
  annotate("text",x=35,y=180000, label="Average Annual Salary: $119563.08 ,",size=3.3 )+
  annotate("text",x=35,y=170000, label="20 states get more than annual average salary)",size=3.3 )
```


```{r}

# Create bar plot
p <- ggplot(data_analyst, aes(x = State, y = Annual_Salary, fill = "Data Analyst", label = State)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = "green", name = "Role", labels = "Data Analyst") +
  labs(title = "Annual Salary for Data Analyst role in US States",
       x = "State", y = "Annual Salary (in dollars)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5, margin = margin(b = 10)))  # Adjust margin
# Set breaks for y-axis
p<-p + scale_y_continuous(breaks = seq(0, max(data_analyst$Annual_Salary), by=10000))

# Add a line for the average annual salary
p <- p + geom_segment(aes(x = 0, xend =50.5, y = avg_salary3, yend = avg_salary3), color = "black", linewidth = 0.001)

#Label average annual salary line
p <- p + annotate("text", x =38, y = avg_salary3+2500, label = paste("Avg Salary Line"), 
                  vjust = -1, hjust = -0.1, color = "black", size = 3)
# Add annotations
p + annotate("text", x=35, y = 140000, label = "(Maximum Annual Salary: $98238 in New York,", size=3.3)+ 
  annotate("text", x =35, y = 130000, label = "Minimum Annual Salary: 62193 in Arkansas ,",size=3.3)+
  annotate("text",x=35,y=120000, label="Average Annual Salary: $75121.02 ,",size=3.3 )+
  annotate("text",x=35,y=110000, label="25 states get more than annual average salary)",size=3.3 )

```

```{r}
# Create bar plot
p <- ggplot(business_analyst, aes(x = State, y = Annual_Salary, fill = "Business Analyst", label = State)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = "sky blue", name = "Role", labels = "Business Analyst") +
  labs(title = "Annual Salary for Business Analyst role in US States",
       x = "State", y = "Annual Salary (in dollars)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5, margin = margin(b = 10)))  # Adjust margin
# Set breaks for y-axis
p<-p + scale_y_continuous(breaks = seq(0, max(business_analyst$Annual_Salary), by=12000))

# Add a line for the average annual salary
p <- p + geom_segment(aes(x = 0, xend =50.5, y = avg_salary4, yend = avg_salary4), color = "black", linewidth = 0.001)


#Label average annual salary line
p <- p + annotate("text", x =36, y = avg_salary4+2500, label = paste("Avg Salary Line"), 
                  vjust = -1, hjust = -0.1, color = "black", size = 3)
# Add annotations
p + annotate("text", x=35, y = 160000, label = "(Maximum Annual Salary: $108915 in Washington,", size=3.3)+ 
  annotate("text", x =35, y = 150000, label = "Minimum Annual Salary: $71249 in Arkansas ,",size=3.3)+
  annotate("text",x=35,y=140000, label="Average Annual Salary: $86956.96 ,",size=3.3 )+
  annotate("text",x=35,y=130000, label="25 states get more than annual average salary)",size=3.3 )

```

```{r}
# Create a bar plot
p <- ggplot(df_new, aes(x = Role, y = Annual_Avg_Salary, fill = Role)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0("$", Annual_Avg_Salary)), position = position_dodge(width = 0.7), vjust = -0.2) +
  labs(title = "Annual Average Salary by Role",
       x = "Role", y = "Annual Average Salary(in dollars)",
       fill = "Role") +
  theme_minimal() +
  theme(axis.text.y = element_text(angle = 0, hjust = 0.5), plot.title = element_text(hjust = 0.5, margin = margin(b = 20))) +
  scale_fill_manual(values = c("Data Scientist" = "red", "Data Engineer" = "orange", "Data Analyst" = "green", "Business Analyst" = "skyblue"))

# Show plot
print(p)
```

## Conclusion:
From the plots above, it is evident that:
*Data Engineers receive the highest annual average salary among U.S. states.
*Data Analysts receive the lowest annual average salary among U.S. states.
*Data Scientists earn the highest annual salary of $145,027 in New York and the lowest of $96,232 in Arkansas.
*Data Engineers earn the highest annual salary of $149,976 in Nevada and the lowest of $96,743 in Florida.
*Data Analysts earn the highest annual salary of $98,238 in New York and the lowest of $62,193 in Arkansas.
*Business Analysts earn the highest annual salary of $108,915 in Washington and the lowest of $71,249 in Arkansas.
In this analysis the specific experience level for each role was unknown. 


## Data References:
1. Data Engineer: https://www.ziprecruiter.com/Salaries/What-Is-the-Average-DATA-Engineer-Salary-by-State
2. Data Analyst: https://www.ziprecruiter.com/Salaries/What-Is-the-Average-Data-Analyst-Salary-by-State
3. Data Scientist: https://www.ziprecruiter.com/Salaries/What-Is-the-Average-DATA-Scientist-Salary-by-State
4. Business Analyst: https://www.ziprecruiter.com/Salaries/What-Is-the-Average-Business-Analyst-Salary-by-State




